---
title: "STT 2820 Test 1 Review"
author: "Alan Arnholt"
date: "2019/02/08 (updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    seal: true
    # lib_dir: libs
    css: [appstate.css, appstate-fonts.css]
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      titleSlideClass: [top, right, inverse]
      # autoplay: 40000
      # countdown: 30000
---

```{r label = "setup", include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA, warning = FALSE, message = FALSE)
```


# Identify the W's for the description of data

- Who
--

- What
--

- When
--

- Where
--

- Why
--

- How

---

# Classify the variable as categorical or quantitative

* Example of categorical variable

--

* Example of quantitative variable

---

# Contingency table

```{r, echo = FALSE}
library(PASWR2)
T1 <- (xtabs(~sample(sex) + sample(pclass), data = TITANIC3))
addmargins(T1)
```

Find the conditional distribution of `sex` for `3rd` class passengers.

--

```{r, echo = FALSE}
A1 <- prop.table(T1, 2)
A1[, 3]
```

--

Find the conditional distribution of `pclass` for `male` passengers.

--

```{r, echo = FALSE}
A2 <- prop.table(T1, 1)
A2[2, ]
```

--

What percent of females are `2nd` class?

--

```{r, echo = FALSE}
prop.table(T1, 1)[1, 2]
#
# 106/(144 + 106 + 216)  only when T1 fixed.
```

---

Based on the density plot below, do you expect the mean or the median to larger?


```{r, echo = FALSE, fig.align = "center", fig.height = 4, fig.width = 4}
library(ggplot2)
ggplot(data = EPIDURALF, aes(x = kg)) +
  geom_density(fill = "goldenrod") + 
  theme_bw() +
  labs(x = "weight of patient in kg.")
```

--

```{r, echo = FALSE}
Mean <- c(mean(EPIDURALF$kg))
Median <- c(median(EPIDURALF$kg))
ans <- c(Mean, Median)
names(ans) <- c("Mean", "Median")
ans
```


---

# Summary Statistics


Last year 10 faculty retired from the College of Arts and Sciences.  Their retirement
ages (in years) are listed below.

```{r, echo = FALSE}
ages <- round(rnorm(10, 65, 2),0)
ages
```

* Find the mean retirement age.

--

The mean retirement age was `r mean(ages)` years.

--

* Find the standard deviation of retirement age.

--

The standard deviation of retirement age was `r round(sd(ages),2)` years.

--

* Find the five-number summary for the retirement ages.

--

The five-number summary is `r fivenum(ages)`.

---

If your friend's mother was `r ages[2]` year old when she retired, how many standard deviations away from the mean was your friend's mother when she retired?

```{r, echo = FALSE}
ages
```

--

```{r, echo = FALSE}
z <- (ages[2] - mean(ages))/sd(ages)
```

Your friend's mother was  `r abs(round(z, 2))` standard deviations `r ifelse(sign(z) < 0, "below", "above")` the average retirement age.

---

# The Standford Binet IQ exam has a mean of 100 and a standard deviation of 16. This is generally written as IQ $\sim N(100, 16)$.

```{r, echo = FALSE}
x <- round(rnorm(1, 110, 5))
```


* What percentage of the population can we expect to have an IQ less than `r x`?

--

```{r, echo = FALSE}
z <- (x - 100)/16
ans <- pnorm(x, 100, 16)
```

$$z = \frac{x - \mu}{\sigma}$$

We can expect `r round(ans*100, 2)`% of the population to have an IQ less than `r x`.

---

# Find the indicated quantities

```{r, echo = FALSE}
zs <- rnorm(5)
```


* $z < `r zs[1]`$
--
```{r, echo = FALSE}
pnorm(zs[1])
```
--

* $z > `r zs[2]`$
--
```{r, echo = FALSE}
pnorm(zs[2], lower = FALSE)
```
--
* $`r min(zs)` < z < `r max(zs)`$
--
```{r, echo = FALSE}
pnorm(max(zs)) - pnorm(min(zs))
```

---

```{r, echo = FALSE}
perc <- sample(2:12, 1)
```


Based on the normal model for male NCAA Div I swim speed for the 100 yards freestyle $N(45, 2)$, what is the cutoff value for the fastest `r perc`% of the swims?

--

```{r, echo = FALSE}
round(qnorm(perc/100, 45, 2),2)
```

---

Find the missing parameter from a normal distributions with the following characteristics:

```{r, echo = FALSE}
val <- rnorm(4, 30, 5)
```



* $\mu = `r round(mean(val),2)`$, 37% below `r round(min(val),2)`; $\sigma=$?

--

```{r, echo = FALSE}
z <- qnorm(.37)
sigma <- (round(min(val),2) - round(mean(val),2))/z
sigma
```
--

* $\sigma = `r round(min(val)/5,2)`$, 20% below `r round(min(val),2)`; $\mu=$?

--

```{r, echo = FALSE}
round(min(val),2) - qnorm(.20)*round(min(val)/5,2)
```

---

# Find the correlation.

The paired data below consist of the test scores of 6 randomly selected students and the number of hours they slept the night before their test.

```{r, echo = FALSE}
sleep <- sort(round(rnorm(6, 6, 2),1)) 
grade <- sort(round(rnorm(6, 80, 8), 1))
DF <- data.frame(sleep = sleep, grade = grade)
knitr::kable(DF, format = "html")
```

--

```{r, echo = FALSE}
cor(sleep, grade)
```

---

```{r, echo = FALSE}
mod <- lm(grade ~ sleep)
```
```{r, echo = FALSE}
nbs <- sample(2:5, 1)
```


* The relationship between grade on a test and the number of hours slept is analyzed.  A regression to predict the average grade from the number of hours slept has an $R^2 = `r round(cor(sleep, grade)^2*100,2)`$%.  The regression equation is $\widehat{grade} = `r coef(mod)[1]` + `r coef(mod)[2]`\cdot sleep$.  Predict the average grade for a student who sleeps `r nbs` hours the night before the exam.

--

```{r, echo = FALSE}
pre <- predict(mod, newdata = data.frame(sleep = nbs))
pre
```
--

* A student slept `r nbs` hours the night before the exam and scored a 72 on the test.  What is the residual for this student?

--

```{r, echo = FALSE}
resid <- 72 - pre
resid
```

