---
title: "STT 2820 Chapter 14"
author: "Alan T. Arnholt"
date: "updated: `r Sys.Date()`"
output:
  xaringan::moon_reader:
    seal: true
    # lib_dir: libs
    css: [appstate.css, appstate-fonts.css]
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      titleSlideClass: [top, right, inverse]
      # autoplay: 40000
      # countdown: 30000
---
class: inverse, center, middle

```{r label = "setup", include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA, warning = FALSE, message = FALSE)
```


# Diagrams

---
National Pet Owners Survey 39% of US households own at least one dog and 34%
of US households own at least one cat.  Assume that 60% of US households own a
dog or cat.  **DRAW A PICTURE**

a.  What is the probability that a randomly selected household owns neither a dog nor a cat?

--

0.4

--

b.  What is the probability that a randomly selected household owns both a dog and a cat?

--

0.13

--

c.  What is the probability that a randomly selected household owns a cat if that household owns a dog?

--

1/3
---
class: inverse, center, middle

# Tree

---
A manufacturing firm orders computer chips from three companies.  10% from
Company A, 20% from B and 70% from C.  Some are defective.  4% of A's are
defective; 2% of Bs are; and 0.5% of C's are.  A worker discovers a randomly
selected chip is defective.  What is the probability it came from Company B?

--

$$\frac{0.20 \times 0.02}{0.10 \times 0.04 +  0.20 \times 0.02 + 0.70 \times 0.005} = `r (.2*.02)/(0.1*.04 + 0.20*.02 + 0.7 * 0.005)`$$

---

# Conditional Probability

We often find probabilities of the form "given B, what is the probability of A?"  These are called **conditional probabilities** and are written $P(A | B) = P(A \cap B)/P(B)$.  If A and B are independent, $P(A | B)= P(A)$, $P(B | A)= P(B)$ and the formula we used for independence follows...ie, $P(A \cap B) = P(A)\cdot P(B)$.

--

```{r, echo = FALSE}
vec <- c(30, 145, 42, 203)
mat <- matrix(vec, nrow = 2)
dimnames(mat) <- list(Laptop = c("Yes", "No"), Searched = c("Yes", "No"))
addmargins(mat)
```

--

Given this information, would you say being searched is independent of carrying
a laptop?

--

$P(\text{laptop} \cap \text{searched}) = \frac{30}{420} = \frac{1}{14}$...the question is does this equal $P(\text{laptop}) \times P(\text{searched}) = \frac{72}{420} \cdot \frac{175}{420} = \frac{1}{14} \Rightarrow$ the events being serached and carrying a laptop are independent.


---
class: inverse, center, middle

# Bayes Rule

---

Ovarian cancer affects only 1 of every 5000 women.  A new test detects this
cancer 99.97% of the time when it is present.  Unfortunately, it also gives a
false positive 5% of the time.  What is the probability that a woman who tests
positive actually has ovarian cancer? **DRAW A TREE**

--

In general $P(A | B) = \dfrac{P(A \cap B)}{P(B)}=\dfrac{P(B | A)\cdot P(A)}{P(B | A) P(A) + P(B | A^c)P(A^c)}$.  

Sometimes $A^c$ is composed of multiple events, all of which must be considered.

--

$P(C|+) = \dfrac{P(C \cap +)}{P(+)}=\dfrac{P(+ | C)\cdot P(C)}{P(+ | C) P(C) + P(+ | C^c)P(C^c)}$

Answer: $\frac{0.9997 \times 0.0002}{0.9997 \times 0.0002 + 0.05 \times 0.9998} = `r 0.0002*0.9997/(0.0002*0.9997 + 0.9998*0.05)`$

---
class: inverse, center, middle

# Sampling Without Replacement

---

# Consider a deck of cards.

a.  What is the probability of getting two aces in the first two cards?

--

$P(A_2 | A_1)\times P(A_1) = 3/51 \times 4/52 = 12/2652$

--

b. What is the probability of getting five hearts in a row?

--

$P(\text{five hearts in a row}) = 13/52 \times 12/51 \times 11/50 \times 10/50 \times 9/49 =\\ `r 13/52*12/51*11/50*10/50*9/49`$.